---
title: 'The metaSEM package'
author: "Mike Cheung"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    self_contained: yes
    theme: united
    toc: yes
    keep_md: true
  pdf_document:
    toc: yes
  word_document: default
---

```{r, echo=FALSE}
library(knitr)
opts_chunk$set(comment=NA, message=FALSE)
```

# Introduction
* The `metaSEM` package provides functions to conducting univariate and multivariate meta-analysis using a structural equation modeling approach via the [OpenMx](http://openmx.psyc.virginia.edu/) package. It also implemented the two-stage structural equation modeling (TSSEM) approach to conducting fixed- and random-effects meta-analytic structural equation modeling (MASEM) on correlation/covariance matrices. 

* The `metaSEM` package is based on the following papers:
    + Cheung, M.W.-L. (2008). A model for integrating fixed-, random-, and mixed-effects meta-analyses into structural equation modeling. _Psychological Methods_, _13_, 182-202.
    + Cheung, M.W.-L. (2009). Constructing approximate confidence intervals for parameters with structural equation models. _Structural Equation Modeling_, _16_, 267–294.
    + Cheung, M.W.-L. (2010). Fixed-effects meta-analyses as multiple-group structural equation models. _Structural Equation Modeling_, _17_, 481-509.
    + Cheung, M.W.-L. (2013). Implementing restricted maximum likelihood estimation in structural equation models. _Structural Equation Modeling_, _20_, 157-167.
    + Cheung, M.W.-L. (2013). Multivariate meta-analysis as structural equation models. _Structural Equation Modeling_, _20_, 429-454. 
    + Cheung, M.W.-L. (2014). MetaSEM: an R package for meta-analysis using structural equation modeling. _Frontiers in Psychology_, _5 (1521)_.
    + Cheung, M.W.-L. (2014). Fixed- and random-effects meta-analytic structural equation modeling: Examples and analyses in R. _Behavior Research Methods_, _46_, 29-40. 
    + Cheung, M.W.-L. (2014). Modeling dependent effect sizes with three-level meta-analyses: A structural equation modeling approach. _Psychological Methods_, _19_, 211-229.
    + Cheung, M.W.-L. (2015). Meta-Analysis: A Structural Equation Modeling Approach. [Wiley](http://as.wiley.com/WileyCDA/WileyTitle/productCd-1119993431.html). 
    + Cheung, M.W.-L., & Chan, W. (2004). Testing dependent correlation coefficients via structural equation modeling. _Organizational Research Methods_, _7_, 206–223.
    + Cheung, M.W.-L., & Chan, W. (2005). Meta-analytic structural equation modeling: A two-stage approach. _Psychological Methods_, _10_, 40-64.
    + Cheung, M.W.-L., & Chan, W. (2009). A two-stage approach to synthesizing covariance matrices in meta-analytic structural equation modeling. _Structural Equation Modeling_, _6_, 28-53.

# Examples
## Univariate meta-analysis
### Random-effects model
```{r}
## Load the library
library(metaSEM)

## Show the first few studies of the data set
head(Becker83)

## Random-effects meta-analysis with ML
summary( meta(y=di, v=vi, data=Becker83) )
```

### Fixed-effects model
```{r}
## Fixed-effects meta-analysis by fixiing the heterogeneity variance at 0
summary( meta(y=di, v=vi, data=Becker83, RE.constraints=0) )
```

### Mixed-effects model
```{r}
## Mixed-effects meta-analysis with "log(items)" as the predictor
summary( meta(y=di, v=vi, x=log(items), data=Becker83) ) 
```

## Multivariate meta-analysis
### Random-effects model
```{r}
## Show the data set
Berkey98

## Multivariate meta-analysis with a random-effects model
mult1 <- meta(y=cbind(PD, AL), v=cbind(var_PD, cov_PD_AL, var_AL), data=Berkey98) 

summary(mult1)

## Plot the effect sizes
plot(mult1)

## Plot the effect sizes with the forest plots
## Load the library for forest plots 
library("metafor")

## Create extra panels for the forest plots
plot(mult1, diag.panel=TRUE, main="Multivariate meta-analysis",
axis.label=c("PD", "AL"))

## Forest plot for PD
forest( rma(yi=PD, vi=var_PD, data=Berkey98) )
title("Forest plot of PD")

## Forest plot for AL
forest( rma(yi=AL, vi=var_AL, data=Berkey98) )
title("Forest plot of AL")
```

### Fixed-effects model
```{r}
## Fixed-effects meta-analysis by fixiing the heterogeneity variance component at 
## a 2x2 matrix of 0.
summary( meta(y=cbind(PD, AL), v=cbind(var_PD, cov_PD_AL, var_AL), data=Berkey98,
              RE.constraints=matrix(0, nrow=2, ncol=2)) )
```

### Mixed-effects model
```{r}
## Multivariate meta-analysis with "publication year-1979" as a predictor 
summary( meta(y=cbind(PD, AL), v=cbind(var_PD, cov_PD_AL, var_AL), data=Berkey98,
              x=scale(pub_year, center=1979)) )
```

## More examples
* [Three-level meta-analysis](./3level.html)
* [Meta-analytic structural equation modeling](./masem.html)

# Installation and help
* The current version is 0.9-2.
* First of all, you need [R](http://www.r-project.org/) to run it. Since `metaSEM` depends on [OpenMx](http://openmx.psyc.virginia.edu/), you may also need to install `OpenMx` by run the following command inside an `R` session:
```
source('http://openmx.psyc.virginia.edu/getOpenMx.R') 
```

## Windows platform
* Download the Windows [Binary for Windows](https://dl.dropboxusercontent.com/u/25182759/metaSEM_0.9-2.zip) package of the `metaSEM`. Please note that d:\ in Windows is represented by either d:/ or d:\\ in `R`. If the file is saved at d:\\, run the following command inside an `R` session:
```
install.packages(pkgs="d:/metaSEM_0.9-2.zip", repos=NULL)
```

## Linux and Mac OS X platforms
* Download the [Source](https://dl.dropboxusercontent.com/u/25182759/metaSEM_0.9-2.tar.gz) package of `metaSEM`. If the file is saved at /home/mike/, run the following command inside an `R` session:
```
install.packages(pkgs="/home/mike/metaSEM_0.9-2.tar.gz", repos=NULL, type="source")
```
* Alternatively, the `metaSEM` package may also be installed in the terminal by using:
```
R CMD INSTALL metaSEM_0.9-2.tar.gz
```

## Help
* [Reference manual](./manual/index.html)
* [Vignette](https://dl.dropboxusercontent.com/u/25182759/metaSEM.pdf)
* If you need help,
  	 + **[OpenMx discussion forum](http://openmx.psyc.virginia.edu/forums/third-party-software/metasem)**: A discussion forum for the `metaSEM` package in `OpenMx`. You may post technical questions related to `metaSEM` there. Please include information on the session. It will be helpful if you can include a reproducible example. You may save a copy of your data, say `my.df`, and attach the content of `myData.R` in the post by using
```
sessionInfo()
dump(c("my.df"), file="myData.R")
```
	+ **[Meta-analysis resources](http://meta-analysis.ning.com/group/meta-analytic-structural-equation-modeling?xg_source=activity)**: A discussion forum related to the conceptual issues of meta-analytic structural equation modeling.

The files are based on the following versions of `R` and `R` packages:
```{r}
sessionInfo()
```
