---
title: 'Examples of MASEM'
author: "Mike Cheung"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    self_contained: yes
    theme: united
    toc: yes
    keep_md: true
  pdf_document:
    toc: yes
  word_document: default
---

```{r, echo=FALSE}
library(knitr)
opts_chunk$set(comment=NA, message=FALSE)
```

# Digman (1997)
## Inspect the data
The correlation matrices and the sample sizes were stored in `Digman97$data` and `Digman97$n`, respectively. We may list the first few cases of the data by using the `head()` command.
```{r}
  #### Load the metaSEM library for TSSEM
  library(metaSEM)
  
  #### Inspect the data for inspection
  head(Digman97$data)

  head(Digman97$n)
```

## Fixed-effects TSSEM
### Stage 1 analysis
To conduct a fixed-effects TSSEM, we may specify `method=FEM` argument (the default method) in calling the `tssem1()` function. The results are stored in an object named `fixed1`. It can be displayed by the `summary()` command. The $\chi^2(130, N=4,496) = 1,499.73, p < .001$, CFI=0.6825, RMSEA=0.1812 and SRMR=0.1750. Based on the test statistic and the goodness-of-fit indices, the assumption of homogeneity of correlation matrices was rejected. 
```{r}
## Fixed-effects model: Stage 1 analysis
fixed1 <- tssem1(my.df=Digman97$data, n=Digman97$n, method="FEM")
summary(fixed1)
```

The pooled correlation matrix (the parameter estimates) can be extracted by the use of the `coef()` command.
```{r}
coef(fixed1)
```

### Stage 2 analysis
As an illustration, I continued to fit the structural model based on the pooled correlation matrix. We may specify the two-factor model with the [RAM formulation](http://openmx.psyc.virginia.edu/docs/OpenMx/latest/Examples_Matrix.html)
```{r}
## Factor covariance among latent factors
Phi <- matrix(c(1,"0.3*cor","0.3*cor",1), ncol=2, nrow=2)

## Error covariance matrix
Psi <- Diag(c("0.2*e1","0.2*e2","0.2*e3","0.2*e4","0.2*e5"))

## S matrix
S1 <- bdiagMat(list(Psi, Phi))

## This step is not necessary but it is useful for inspecting the model.
dimnames(S1)[[1]] <- dimnames(S1)[[2]] <- c("A","C","ES","E","I","Alpha","Beta") 

## Display S1
S1

## A matrix
Lambda <-
matrix(c(".3*Alpha_A",".3*Alpha_C",".3*Alpha_ES",rep(0,5),".3*Beta_E",".3*Beta_I"),
       ncol=2, nrow=5)
A1 <- rbind( cbind(matrix(0,ncol=5,nrow=5), Lambda),
             matrix(0, ncol=7, nrow=2) )

## This step is not necessary but it is useful for inspecting the model.
dimnames(A1)[[1]] <- dimnames(A1)[[2]] <- c("A","C","ES","E","I","Alpha","Beta") 

## Display A1
A1

## F matrix to select the observed variables
F1 <- create.Fmatrix(c(1,1,1,1,1,0,0), as.mxMatrix=FALSE)

## Display F1
F1
```
 
Since we are conducting a correlation structure analysis, the error variances are not free parameters. We need to use the `diag.constraints` to constrain the diagonals as one. When there are nonlinear constraints, standard errors are not available in `OpenMx`. We may request the likelihood-based confidence intervals (LBCI) with the `intervals` argument.
The chi-square test statistic of the Stage 2 analysis was $\chi^2(4, N=4,496) = 65.06, p < .001$, CFI=0.9802, RMSEA=0.0583 and SRMR=0.0284.
```{r}
fixed2 <- tssem2(fixed1, Amatrix=A1, Smatrix=S1, Fmatrix=F1, 
                 diag.constraints=TRUE, intervals="LB",
                 model.name="TSSEM2 Digman97")
summary(fixed2)
```

## Fixed-effects TSSEM with cluster
### Stage 1 analysis
There are 4 types of sample characteristics in the original `cluster`. We may group them into either younger or older samples. 
```{r}
#### Display the frequencies of "cluster"
table(Digman97$cluster)

#### Fixed-effects TSSEM with several clusters
#### Create a variable for different cluster
#### Younger participants: Children and Adolescents
#### Older participants: others
clusters <- ifelse(Digman97$cluster %in% c("Children","Adolescents"),
                   yes="Younger participants", no="Older participants")
  
#### Show the clusters
clusters
```

We may conduct a fixed-effects model by specifying the `cluster=clusters` argument. Fixed-effects TSSEM will be conducted according to the labels in the `clusters`. The goodness-of-fit indices of the Stage 1 analysis for the older and younger participants were $\chi^2(80, N=3,658) = 823.88, p < .001$, CFI=0.7437, RMSEA=0.1513 and SRMR=0.1528, and $\chi^2(40, N=838) = 344.18, p < .001$, CFI=0.7845, RMSEA=0.2131 and SRMR=0.1508, respectively. 
```{r}
## Example of Fixed-effects TSSEM with several clusters
cluster1 <- tssem1(Digman97$data, Digman97$n, method="FEM", 
                   cluster=clusters)
summary(cluster1)
```

The pooled correlation matrix for each cluster can be extracted by the use of the =coef()= command.
```{r}
coef(cluster1)
```

### Stage 2 analysis
The goodness-of-fit indices of the Stage 2 analysis for the older and younger participants were $\chi^2(4, N=3,658) = 21.92, p < .001$, CFI=0.9921, RMSEA=0.0350 and SRMR=0.0160, and $\chi^2(4, N=838) = 144.87, p < .001$, CFI=0.9427, RMSEA=0.2051 and SRMR=0.1051, respectively. 
```{r}
cluster2 <- tssem2(cluster1, Amatrix=A1, Smatrix=S1, Fmatrix=F1, 
                   diag.constraints=TRUE, intervals.type="z")
#### Please note that the estimates for the younger participants are problematic.
summary(cluster2)
```

We may also plot the parameter estimates of these two groups.
```{r}
library(semPlot)

## Convert the model to semPlotModel object with 2 plots
my.plots <- lapply(X=cluster2, FUN=meta2semPlot, latNames=c("Alpha","Beta"))

## Setup two plots
layout(t(1:2))
semPaths(my.plots[[1]], whatLabels="est", nCharNodes=10, color="green")
title("Younger") 
semPaths(my.plots[[2]], whatLabels="est", nCharNodes=10, color="yellow")
title("Older")
```

## Random-effects TSSEM
### Stage 1 analysis
Since there is not enough data to estimate the full variance component of the random effects, I estimate the variance component with a diagonal matrix in the `RE.type="Diag"` argument. The range of $I^2$ indices, the percentage of total variance that can be explained by the between study effect, are from .84 to .95. 
```{r}  
#### Random-effects TSSEM with random effects on the diagonals
random1 <- tssem1(Digman97$data, Digman97$n, method="REM",
                  RE.type="Diag")
summary(random1)
```

The pooled correlation coefficients (fixed effects) and the variance components (the random effects) can be extracted by the use of the `coef()` command via the `select` argument.
```{r}
## Fixed effects
coef(random1, select="fixed")

## Random effects
coef(random1, select="random")
```

### Stage 2 analysis
The steps are exactly the same as those in the fixed-effects model. The chi-square test statistic of the Stage 2 analysis was $\chi^2(4, N=4,496) = 8.51, p < .001$, CFI=0.9911, RMSEA=0.0158 and SRMR=0.0463. 
```{r}
random2 <- tssem2(random1, Amatrix=A1, Smatrix=S1, Fmatrix=F1, 
                  diag.constraints=TRUE, intervals="LB")
summary(random2)
```

We may also plot the models with the labels or the parameter estimates.
```{r}
library(semPlot)

## Convert the model to semPlotModel object
my.plot <- meta2semPlot(random2, latNames=c("Alpha","Beta"))

## Plot the model with labels
semPaths(my.plot, whatLabels="path", nCharEdges=10, nCharNodes=10, color="red")

## Plot the parameter estimates
semPaths(my.plot, whatLabels="est", nCharNodes=10, color="green")
```

# Becker and Schram (1994)
## Inspect the data
We may list the first few cases of the data by using the `head()` command.
```{r}
#### Load the metaSEM library for TSSEM
library(metaSEM)
  
#### Inspect the data for inspection (not required for the analysis)
head(Becker94$data)
 
head(Becker94$n)
```

## Fixed-effects TSSEM
### Stage 1 analysis
The test statistic of the Stage 1 analysis based on the TSSEM approach was $\chi^2(27, N=538) = 62.50, p < .001$, CFI=0.7943, RMSEA=0.1565 and SRMR=0.2011. Based on the test statistic and the goodness-of-fit indices, the hypothesis of homogeneity of correlation matrices was rejected. 
```{r}
#### Fixed-effects model
## Stage 1 analysis
fixed1 <- tssem1(Becker94$data, Becker94$n, method="FEM")
summary(fixed1)

coef(fixed1)
```

### Stage 2 analysis
We may specify the model via the RAM formulation. If all variables are observed, there is no need to specify the *F* matrix. Since the *df* of the regression model is 0, the proposed model always fits the data perfectly.
```{r}
#### Prepare models for stage 2 analysis
## A1: asymmetric matrix (regression coefficients)
A1 <- create.mxMatrix(c(0,0,0,"0.2*Spatial2Math",
                        0,0,"0.2*Verbal2Math",0,0), 
                        type="Full", ncol=3, nrow=3, 
                        name="A1")
A1 

## S1: symmetric matrix (variance covariance matrix among variables)
S1 <- create.mxMatrix(c("0.2*ErrorVarMath",0,0,1,"0.2*CorBetweenSpatialVerbal",1), 
                        type="Symm", name="S1")
S1

## Stage 2 analysis
fixed2 <- tssem2(fixed1, Amatrix=A1, Smatrix=S1, diag.constraints=TRUE, 
                 intervals="LB")
summary(fixed2)
```

## Fixed-effects TSSEM with cluster
### Stage 1 analysis
The goodness-of-fit indices of the Stage 1 analysis for the female and male participants were $\chi^2(12, N=235) = 42.41, p < .001$, CFI=0.7116, RMSEA=0.2327 and SRMR=0.2339, and $\chi^2(12, N=303) = 16.13, p = .1852$, CFI=0.9385, RMSEA=0.0755 and SRMR=0.1508, respectively.
```{r}
#### Fixed-effects model with cluster
## Stage 1 analysis
cluster1 <- tssem1(Becker94$data, Becker94$n, method="FEM", cluster=Becker94$gender)
summary(cluster1)

coef(cluster1)
```

### Stage 2 analysis
```{r}
## Stage 2 analysis
cluster2 <- tssem2(cluster1, Amatrix=A1, Smatrix=S1, diag.constraints=TRUE, 
                   intervals="LB")
summary(cluster2)
```

We may also plot the parameter estimates of these two groups.
```{r}
## Convert the model to semPlotModel object with 2 plots
my.plots <- lapply(X=cluster2, FUN=meta2semPlot)

## Setup two plots
layout(t(1:2))
semPaths(my.plots[[1]], whatLabels="est", nCharNodes=10, color="green")
title("Females") 
semPaths(my.plots[[2]], whatLabels="est", nCharNodes=10, color="yellow")
title("Males")
```

## Random-effects TSSEM
### Stage 1 analysis
The $I^2$ indices for the correlations between spatial and math, verbal and math, and spatial and verbal are .00, .81 and .23, respectively. 
```{r}
#### Random-effects model
## Stage 1 analysis: A diagonal matrix for random effects 
random1 <- tssem1(Becker94$data, Becker94$n, method="REM", RE.type="Diag")
summary(random1)

coef(random1, select="fixed")

coef(random1, select="random")
```

### Stage 2 analysis
```{r}
## Stage 2 analysis
random2 <- tssem2(random1, Amatrix=A1, Smatrix=S1, diag.constraints=TRUE, 
                  intervals="LB")
summary(random2)
```

We may also plot the models with the labels or the parameter estimates.
```{r}
## Convert the model to semPlotModel object
my.plot <- meta2semPlot(random2)

## Plot the model with labels
semPaths(my.plot, whatLabels="path", nCharEdges=10, nCharNodes=10, color="red")

## Plot the parameter estimates
semPaths(my.plot, whatLabels="est", nCharNodes=10, color="green")
```
